### 2016-08-25 ms_nihao_budget.tex
first version of the draft for the second referee's comments

\documentclass[useAMS,usenatbib]{mn2e}
\input psfig.sty
\usepackage{array}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{graphicx}
%\usepackage{epstopdf}

\voffset=-0.8in

\def \gas {\textsc{gasoline}}
\def \gastwo {\textsc{ESF-gasoline2}}
\def \changa {\textsc{changa}}
\def \mean#1{\left< #1 \right>}

%%MISC
\def \aj {AJ}
\def \apj {ApJ}
\def \apjl {ApJL}
\def \mnras {MNRAS}
\def \apjs {ApJS}
\def \aap {A\&A}
\def \nat {Nature}
\def \pasp {PASP}
\def \na {NewA}

\def \etal {et~al.~}
\def \eg{e.g.}
\def \Section{\S}
\def \spose#1{\hbox  to 0pt{#1\hss}}  
\def \lta{\mathrel{\spose{\lower 3pt\hbox{$\sim$}}\raise  2.0pt\hbox{$<$}}}
\def \gta{\mathrel{\spose{\lower  3pt\hbox{$\sim$}}\raise 2.0pt\hbox{$>$}}}
\def \ion#1#2{#1{\footnotesize{#2}}\relax}
\def \ha       {H$\alpha$}
\def \hi       {\ion{H}{I}}
\def \hii      {\ion{H}{II}}
\def \oii      {[\ion{O}{II}]}
\def \ovi      {\ion{O}{VI}}

%%UNITS
\def \kmsmpc {\>{\rm km}\,{\rm s}^{-1}\,{\rm Mpc}^{-1}}
\def \kms {\ifmmode  \,\rm km\,s^{-1} \else $\,\rm km\,s^{-1}  $ \fi }
\def \kpc {\ifmmode  {\rm kpc}  \else ${\rm  kpc}$ \fi  }  
\def \hkpc {\ifmmode  {h^{-1}\rm kpc}  \else ${h^{-1}\rm kpc}$ \fi  }  
\def \hMpc {\ifmmode  {h^{-1}\rm Mpc}  \else ${h^{-1}\rm Mpc}$ \fi  }  
\def \Msun {\ifmmode \rm M_{\odot} \else $\rm M_{\odot}$ \fi}
\def \hMsun {\ifmmode h^{-1}\,\rm M_{\odot} \else $h^{-1}\,\rm M_{\odot}$ \fi}
\def \hhMsun {\ifmmode h^{-2}\,\rm M_{\odot}\else $h^{-2}\,\rm M_{\odot}$ \fi}
\def \Lsun {\ifmmode L_{\odot} \else $L_{\odot}$ \fi} 
\def \hhLsun {\ifmmode h^{-2}\,\rm L_{\odot} \else $h^{-2}\,\rm L_{\odot}$ \fi}

%%COSMOLOGY

\def\LCDM{$\Lambda$CDM }
\def \LCDM {\ifmmode \Lambda{\rm CDM} \else $\Lambda{\rm CDM}$ \fi}
\def \sig8 {\ifmmode \sigma_8 \else $\sigma_8$ \fi} 
\def \Omegam {\ifmmode \Omega_{\rm m} \else $\Omega_{\rm m}$ \fi} 
\def \Omegab {\ifmmode \Omega_{\rm b} \else $\Omega_{\rm b}$ \fi} 
\def \Omegar {\ifmmode \Omega_{\rm r} \else $\Omega_{\rm r}$ \fi} 
\def \fbar {\ifmmode f_{\rm b} \else $f_{\rm b}$ \fi} 
\def \OmegaL {\ifmmode \Omega_{\rm \Lambda} \else $\Omega_{\rm \Lambda}$\fi} 
\def \Deltavir {\ifmmode \Delta_{\rm vir} \else $\Delta_{\rm vir}$ \fi}
\def \rhocrit {\ifmmode \rho_{\rm crit} \else $\rho_{\rm crit}$ \fi}

%DARK MATTER
\def \rs {\ifmmode r_{\rm s} \else $r_{\rm s}$ \fi} 
\def \rh {\ifmmode r_{\rm h} \else $r_{\rm h}$ \fi} 
\def \Rvir {\ifmmode R_{\rm vir} \else $R_{\rm vir}$ \fi}
\def \Vvir {\ifmmode V_{\rm  vir} \else  $V_{\rm vir}$  \fi} 
\def \Vmax {\ifmmode V_{\rm  max} \else  $V_{\rm max}$  \fi} 
\def \Mvir {\ifmmode M_{\rm  vir} \else $M_{\rm  vir}$ \fi}  
\def \Mhalo {\ifmmode M_{200} \else $M_{200}$ \fi}  
\def \Nvir {\ifmmode N_{\rm  vir} \else $N_{\rm  vir}$ \fi}  
\def \Jvir {\ifmmode J_{\rm vir} \else $J_{\rm vir}$ \fi} 
\def \Evir {\ifmmode E_{\rm vir} \else $E_{\rm vir}$ \fi} 
\def \lam {\ifmmode \lambda  \else $\lambda$ \fi} 
\def \lamp {\ifmmode \lambda^{\prime} \else $\lambda^{\prime}$  \fi} 
\def \lampc {\ifmmode \lambda^{\prime}_{\rm c} \else
  $\lambda^{\prime}_{\rm c}$  \fi}

\def \xoff {\ifmmode x_{\rm off} \else $x_{\rm off}$ \fi}
\def \rhorms {\ifmmode \rho_{\rm rms} \else $\rho_{\rm rms}$ \fi}
\def \qbar {\ifmmode \bar{q} \else $\bar{q}$ \fi}

%%BARYONS
\def \Mb {\ifmmode M_{\rm b} \else $M_{\rm b}$ \fi} 
\def \eSF {\ifmmode \epsilon_{\rm SF} \else $\epsilon_{\rm SF}$ \fi} 
\def \Md {\ifmmode M_{\rm d} \else $M_{\rm d}$ \fi} 
\def \Mg {\ifmmode M_{\rm g} \else $M_{\rm g}$ \fi} 
\def \Rb {\ifmmode R_{\rm b} \else $R_{\rm b}$ \fi} 
\def \Rd {\ifmmode R_{\rm d} \else $R_{\rm d}$ \fi} 
\def \Rg {\ifmmode R_{\rm g} \else $R_{\rm g}$ \fi} 
\def \mgal {\ifmmode m_{\rm gal} \else $m_{\rm gal}$ \fi} 
\def \rj {\ifmmode {\cal R}_j \else ${\cal R}_j$ \fi} 
\def \lamgal {\ifmmode \lambda_{\rm gal} \else $\lambda_{\rm gal}$ \fi} 
\def \Vcirc {\ifmmode V_{\rm circ} \else $V_{\rm circ}$ \fi} 
\def \Vrot {\ifmmode V_{\rm rot} \else $V_{\rm rot}$ \fi} 
\def \Vflat {\ifmmode V_{\rm flat} \else $V_{\rm flat}$ \fi} 
\def \Mstar {\ifmmode M_{\rm star} \else $M_{\rm star}$ \fi} 
\def \Mgas {\ifmmode M_{\rm gas} \else $M_{\rm gas}$ \fi} 
\def \Mbar {\ifmmode M_{\rm bar} \else $M_{\rm bar}$ \fi}
\def \Rbar {\ifmmode R_{\rm bar} \else $R_{\rm bar}$ \fi}

%%MASS-TO-LIGHT RATIOS
\def \DeltaIMF {\ifmmode \Delta_{\rm IMF} \else $\Delta_{\rm IMF}$ \fi}

\def \VV {\ifmmode V_{\rm 2.2}/V_{200} \else $V_{2.2}/V_{200}$ \fi} 
\def \dvr {\ifmmode \partial_{\rm VR} \else $\partial_{\rm VR}$ \fi}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title[Baryon Budget] {NIHAO VII: Predictions for the galactic baryon budget in dwarf to Milky Way mass haloes}

\author[Wang et al.]{Liang Wang$^{1,3}$\thanks{wangliang@pmo.ac.cn}, Aaron A. Dutton$^{2,3}$,
  Gregory S. Stinson$^3$, Andrea V. Macci\`o$^{2,3}$, 
\newauthor{Thales Gutcke$^3$, Xi Kang$^1$}\\
$^1$Purple Mountain Observatory, the Partner Group of MPI f\"ur Astronomie, 2 West Beijing Road, Nanjing 210008, China\\
$^2$New York University Abu Dhabi, PO Box 129188, Abu Dhabi, UAE\\
$^3$Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, 69117 Heidelberg, Germany}
\begin{document}

\date{to be submitted to MNRAS}

\pagerange{\pageref{firstpage}--\pageref{lastpage}}\pubyear{2015}

\maketitle

\label{firstpage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}
  We use the NIHAO galaxy formation simulations to make predictions
  for the baryonic budget  in present day galaxies ranging from dwarf
  ($\Mhalo\sim10^{10} \Msun$) to Milky Way ($\Mhalo\sim10^{12} \Msun$)
  masses.  The sample is made of 88 independent high resolution
  cosmological zoom-in simulations.  NIHAO galaxies reproduce key
  properties of observed galaxies, such as the stellar mass vs halo
  mass and cold gas vs stellar mass relations. Thus they make
  plausible predictions for the baryon budget.  We present the mass
  fractions of stars, cold gas ($T<10^4$K), cool gas ($10^4 < T <
  10^5$K), warm-hot gas ($10^5 < T < 10^7$K), and hot gas (T$> 10^7$K) 
  , inside the virial radius, $R_{200}$.  Compared to the predicted
  baryon mass, using the dark halo mass and the universal baryon
  fraction, $f_{\rm b}\equiv \Omega_{\rm b}/\Omega_{\rm m}=0.15$, we
  find that all of our haloes are missing baryons. The missing
    mass been relocated past 2 virial radii, and {\bf cool gas dominates
  corona at low mass (M$_{200} \lta 3 \times 10^{11} \Msun$) while the 
  warm-hot gas dominates dominates at high mass (M$_{200} \gta 3 \times 10^{11} \Msun$)}.
   Haloes of mass $\Mhalo\sim 10^{10}\Msun$ are missing
  $\sim 90\%$ of their baryons.  More massive haloes ($\Mhalo\sim
  10^{12}\Msun$) retain a higher fraction of their baryons, with $\sim
  30\%$ missing, consistent with recent observational estimates.
  {\bf Moreover, these more massive haloes reproduce the observed 
  fraction of cold, warm-hot and hot gas, the fraction of cool gas we
  predict is significant lower than the observation from COS-HALOs,
  but agree with the analysis of \citet{Stern16}}.
\end{abstract}

\begin{keywords}
  galaxies: evolution -- galaxies: formation -- galaxies: dwarf -- galaxies: spiral -- 
  methods: numerical -- cosmology: theory
\end{keywords}

\setcounter{footnote}{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% SECTION 1: INTRODUCTION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
\label{sec:intro}
Cosmic structure formation has redistributed the baryons from  a
nearly uniformly distributed plasma into a variety of states,
including stars, stellar remnants, cold (atomic and molecular) gas,
and hot (ionized) gas. The theories of galaxy formation can predict
the amount of mass in these different states, which can in turn be tested by
observational constraints.

%% missing baryon problem
On cosmological scales, the ratio between the total baryonic and
gravitating mass is measured to be $f_{\rm b}\equiv \Omega_{\rm
  b}/\Omega_{\rm m}\simeq 0.15$ (The Planck Collaboration 2014).
However, the cold baryonic mass density implied by several galaxy baryon
estimates {\bf (mainly about stars and cold gas)} is only 3-8\% of 
the big bang nucleosynthesis expectation
\citep{Persic92, Fukugita98,  Bell03, McGaugh10}.  The majority of the
cosmic baryons are thought to be in the form of hot gas around or
between galaxies \citep{Cen09}. Until recently only a fraction of
these baryons had been detected \citep{Bregman07, Shull12}.  This
discrepancy is referred to as the ``missing baryon problem''.
Several theoretical studies with cosmological simulations
have constrained the phase of
the potential reservoirs of the missing baryons {\bf in intergalactic
medium (IGM), and find a large fraction of the baryons with low density
and high temperature resides between the galaxies}
\citep{Yoshida05, He05, Dave10, Zhu11, Haider16}.

The circum galactic medium (CGM) is {\bf another}  major potential  reservoir
of the missing baryons. \hi{} and metal absorption lines are
expected  to signpost such diffuse baryonic content. \hi{} is mainly
from gas with temperature T $\sim 10^4$ K so that it is able to
detect cold gas in the CGM.  Meanwhile, theoretical work has
predicted that a substantial  portion of the CGM is in the warm or
hot phase with temperature T $> 10^{4.5}$ K.  Gas enters this phase
through photoionization, accretion shocks or shocks caused by
galactic winds \citep{Voort12}. Such dilute halo gas is at T$\sim
10^{4.5-7}$ K, so the detection is dominated by metal lines,
e.g. \ovi{}.  Recent advances in the detection of gas in the CGM
have come  from the COS survey \citep{Tumlinson11, Tumlinson13,
Thom12,  Werk12, Werk13}.  On the scale of Milky Way mass haloes
$\Mhalo \sim 10^{12}\Msun$ a significant amount of warm ($10^4 < T <
10^7$K) gas has been detected \citep{Werk14}, accounting for 33-88\%
of the baryon budget. In the future such observations will be extended
to a wider range of halo masses.  A number of large volume
cosmological simulations \citep{Ford13, Ford15, Suresh15b,
  Oppenheimer16} and zoom-in cosmological simulations
\citep{Stinson12, Hummels13, Shull14} have given 
predictions for the
\hi{} and \ovi{} absorption line. \citet{Gutcke16} compared the
column density profile of \ovi{} and \hi{} in the CGM of galaxies
from the NIHAO \citep{Wang15} cosmological hydrodynamical
simulation suite with observations, studied the covering fraction of
dense \hi{}, looked at the shape of the CGM and its chemical
composition. The conclusion shows {\bf the simulation can reproduce 
the observational covering fraction and column density profile
of cool \hi{} well, and echo observed \ovi{} column density trend.
As commonly in other simulations, however, the column density of \ovi{}
is lower and the extent of optically thick \hi{} is smaller than
observed.}

The physical properties of the CGM has been shown to be able to
test feedback models \citep{Sharma12,Marasco13}.  \citet{Dave09}
predicted galactic halo baryon fractions of galaxies with halo
masses ranging from $10^{11} \Msun$ to $10^{13} \Msun$ using
cosmological hydrodynamical simulations with a well-constrained
model for galactic outflows.  They found that, without the outflow
model, the baryon fraction inside the virial radius is roughly the
cosmic baryonic fraction, but with the outflow model, the baryon
fraction is increasingly suppressed in lower mass haloes.  By
comparing results at $z=3$ and $z=0$, they showed that large
haloes remove their baryons at early times while small haloes lose
baryons more recently due to the wind material taking longer to
return to low-mass galaxies than high-mass galaxies.
{\bf \citet{Muratov15} showed similar results that the gas and
baryon fractions are lower at lower redshift, after powerful
outflows at intermediate redshift z $\approx 0.5-2$ remove a large
amount of gas from the halo. \citet{Voort16} found baryon fraction
and gas fraction are reduced compared to the cosmic mean baryon
fraction, and low mass haloes reduce larger since feedback can
easily expel gas from it, which is consistent with existing 
observation.}
%  
\citet{Sokolowska16} studied the halo gas of three Milky way-sized
galaxies using cosmological zoom-in simulations. They found that
most of missing baryons actually resides in warm-hot and hot gas
which contribute to 80\% of the total gas reservoir.  The recovered
baryon fraction within 3 virial radii is 90\%.  The warm-hot medium
is sensitive to the feedback model so that a reliable spatial
mapping of the warm-hot medium will provide  a stringent test for
feedback models.

In this paper we make predictions for the baryonic budget for stars,
cold, warm and hot gas in and around the virial radius of haloes of
mass ranging from $\Mhalo\sim 10^{10}\Msun$ to $10^{12}\Msun$. We use
a sample of 88 zoom-in galaxy formation simulations from the NIHAO
project.
Reproducing the stellar mass content in dark matter haloes both
today and in the past has been a formidable challenge for cosmological galaxy
formation simulations \citep{Weinmann12, Hopkins14}. Even the latest
state-of-the art simulations have trouble: the ILLUSTRIS simulation
\citep{Vogelsberger14} strongly overpredicts the stellar masses in
dwarf galaxy haloes $(M_{200} \lta 10^{11}\Msun)$, while the EAGLE
simulations \citep{Schaye15} underpredict the peak of the star
formation efficiency in halos of mass $M_{200}\sim 10^{12}\Msun$. In
contrast, the NIHAO galaxies are consistent with the stellar mass
vs halo mass relations from halo abundance matching since redshift
$z\sim 4$ \citep{Wang15}, the galaxy star formation rate vs stellar
mass relation since $z\sim 4$ \citep{Wang15}, and the cold gas mass vs
stellar mass relation at $z\sim 0$ \citep{Stinson15}.  Therefore, the
simulations make plausible predictions for the mass fractions and
physical locations of the warm and hot gas components.  We find that
all the haloes contain less baryons than expected according to the
cosmic baryonic fraction, but the missing fraction is strongly mass
dependent.

This paper is organized as follows: The cosmological hydrodynamical
simulations including star formation and feedback are briefly
described in  \S\ref{sec:sims}; In \S\ref{sec:budget} we present the
results including the baryonic budget, baryon distribution, and a
comparison with observations; \S\ref{sec:sum} gives a summary of our
results.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% SECTION 2 SIMULATIONS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Simulations} 
\label{sec:sims}

In this study we use simulations from the NIHAO (Numerical
Investigation of a Hundred Astrophysical Objects) project \citep{Wang15}.  
The initial conditions are created to keep the same numerical
resolution across the whole mass range with typically a million dark
matter particles inside the virial radius of the target halo at 
redshift $z=0$.  The halos to be re-simulated at higher resolution
with baryons have been extracted from 3 different pure N-body
simulations with a box size of 60, 20 and 15 $h^{-1}$ Mpc
respectively.  We adopted the  latest compilation of cosmological
parameters from the Planck  satellite \citep{Planck14}. 
More information on the collisionless parent simulations,
{\bf the force softenings and particle masses for the highest 
refinement level for each simulation} and
sample  selection can be found in \citet{Dutton14} and \citet{Wang15}

We use the SPH hydrodynamics code {\sc gasoline} \citep{Wadsley04},
with a revised treatment of  hydrodynamics as described in
\citet{Keller14}.  The code includes a subgrid model for turbulent
mixing of metal and energy \citep{Wadsley08}, heating and cooling
include photoelectric heating of dust grains, ultraviolet (UV) heating
and ionization and  cooling due to hydrogen, helium and metals
\citep{Shen10}.

khe star formation and feedback modeling follows what
was used in the MaGICC simulations \citep{Stinson13}.  
{\bf The gas converts into stars according to the Kennicutt-Schmidt Law
when it satisfies a temperature and density threshold. Stars feed both
metals and energy back into the ISM gas surrounding the region where
they formed. SN feedback is implemented using the blastwave formalism
described in \citet{Stinson06}. Pre-SN feedback is an effort to consider
radiation energy from massive stars which is am important complementary
to physcial model for stellar feedback. Heating is introduced immediately
after massive stars form based on how much star light is radiated. 
Our simulations use thermal feedback to provide pressure support and 
increase gas temperature above the star formation threshold to
decrease star formation.}
There are two
small changes in NIHAO simulations: The change in  number of neighbors
and the new combination of softening length and  particle mass increases
the threshold for star formation from  9.3 to 10.3
cm$^{-3}$, the increase of pre-SN feedback efficiency $\epsilon_{\rm
  ESF}$, from 0.1 to 0.13.  The more detail on star formation and
feedback modeling can be found in \citet{Wang15}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% SECTION 3 RESULTS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Results}
%\label{sec:results}

%% FIGURE 1
\begin{figure}
\centerline{
%  \psfig{figure=barbudget_linear_v3.eps,width=0.5\textwidth}}
%\centerline{
%  \psfig{figure=barbudget_v3.eps,width=0.5\textwidth}
  \psfig{figure=barbudget_linearlog.eps,width=0.48\textwidth}
}
\caption{Fractional baryon content of our NIHAO simulations  as a
  function of halo mass. The green points show the ratio between the
  baryonic mass (stars + gas) inside the virial radius and the total
  baryonic mass associated with the dark matter halo. The blue points
  show the corresponding fraction for the stars. The solid green line
  and shaded region shows a double power-law fit, together with the
  1$\sigma$ scatter. For the stellar mass fraction we show several
  relations from halo abundance matching.    The linear (upper panel)
  and logarithmic (lower panel) scales emphasize the large amount of
  ``missing'' baryons, and the power-law nature of the relations,
  respectively.}
\label{fig:budget}
\end{figure}

%% FIGURE 2
\begin{figure}
\centerline{
  \psfig{figure=m500c_frac.eps,width=0.48\textwidth}
}
\caption{{\bf Baryon fraction within R$_{500}$ comparing with \citet{Voort16}.
   The blue points are from NIHAO simulations and the red pentagons are from
   FIRE simulations.}}
\label{fig:fire_com}
\end{figure}

%% FIGURE 3
\begin{figure}
\centerline{
  \psfig{figure=budget_res.eps,width=0.48\textwidth}
}
\caption{{\bf Linear fractional baryon content of our NIHAO simulations  as a
  function of halo mass color coded by number of all particles per halo.}}
\label{fig:budget_res}
\end{figure}

\section{Baryon budget}
\label{sec:budget}

We define the fiducial baryonic mass as:
\begin{equation}
M_{\rm b} \equiv M_{\rm b}(R_{200})= \frac{f_{\rm b}}{1-f_{\rm b}}M_{\rm dm}(R_{200}) 
\label{equ:mb}
\end{equation}
where the $M_{\rm dm}$ is the total dark matter mass of the halo, and
the $f_{\rm b} = \Omegab/\Omegam \sim 0.15$ is the cosmic baryon
fraction (the ratio between baryon density and mass density including baryonic mass
plus dark matter), so that $M_{\rm b}$ would be the baryonic mass inside the
virial radius if the baryons followed the dark matter closely.

Fig.~\ref{fig:budget} shows the ratio between the  mass of each baryon
component inside the virial radius  to the fiducial baryonic mass for
the most massive galaxy in each zoom-in region. We present the
fractions of total stellar mass (blue points), and the total baryonic
mass including stellar mass plus gas mass (green points).  For the
stellar mass fraction we also show the relations from the halo
abundance  matching \citep{Moster13, Behroozi13, Kravtsov14}.  The
grey area is the one sigma scatter around the mean value  from
\citet{Kravtsov14}.
Fig.~\ref{fig:budget} shows that all haloes in our study contain less
than the universal fraction of baryons. The upper panel uses a linear
y-axis scale, which highlights the large amount of baryons that are
missing, especially in low mass haloes. The logarithmic scale in the
lower panel highlights the power-law nature of the relations.

The trends of each component fraction are similar, in that  the
fractions are relatively low in low mass haloes, and increase as the
halo mass increases.  The main difference between the different
components is the slope, with the baryonic mass fraction having a
shallower slope than the stellar mass fraction.  This is because in
low mass haloes ($\Mhalo\sim 10^{10}\Msun$) most of the baryons are in
the form of gas, while in the highest mass haloes we study
($\Mhalo\sim 10^{12}\Msun$) {\bf a substantial amounts of gas
forms stars}.

{\bf \citet{Christensen16} run twenty galaxy simulations spanning the mass range
from $10^{9.5} - 10^{12} \Msun$ using zoom-in cosmological hydrodynamic 
simulation code and didn't only reproduce the stellar mass - halo mass
relation but also showed a similar baryon fraction relation, e.g. in 
L$^*$ galaxies, the fraction is reduced by 20\% while in galaxies with
halo mass $< 10^{10} \Msun$ there can be as large as an 80\% mass reduction.
This suggests that the potential well of the halo regulates the baryon
fraction.
To compare with baryon fraction in \citet{Voort16}, we derive
the M$_{500}$ and R$_{500}$ to recalculate the ratio between 
baryon fraction to the universal fraction of baryons. The baryon
fraction have good agreement with each other in the whole mass range
where NIHAO simulations explored. For the massive haloes with 
M$_{500} > 10^{12} \Msun$, there are two galaxies from NIHAO have
reduced baryon fraction and four more massive galaxies from FIRE
have larger reduced baryon fraction which may hint that the baryon
fraction will decrease as the halo mass increases in large mass range.}

Since most of the haloes we study are above the mass where the cosmic UV
background prevents gas from cooling, the missing baryons have most
likely been ejected from the central galaxies in supernova/stellar
feedback driven winds.  Although the lower mass galaxies have
converted a smaller fraction of their available baryons into stars,
and hence there is proportionally less energy available to drive an
outflow, they have expelled a larger fraction of their baryons,
consistent with expectations from energy driven gas outflows
\citep[e.g.,][]{Dutton12}.

The behavior of the baryonic mass fraction, $f_{\rm bar}$, as a 
function of the halo mass is captured using a double power law formula:
\begin{equation}
\frac{f}{f_0} = \left( 
                        \frac{M_{200}}{\mathcal{M}_0} 
                        \right)^\alpha 
                        \left\{ 0.5 \left[ 1+\left( 
                        \frac{M_{200}}{\mathcal{M}_0} 
                        \right)^\gamma \right] 
                        \right\}^{\frac{\beta-\alpha}{\gamma}}.
\end{equation}
In this formula, the lower and higher mass ends have logarithmic slope
$\alpha$ and $\beta$, respectively, while $\gamma$ regulates how 
sharp the transition is from the lower to the higher ends.
The best fit parameters are as follows:
\begin{eqnarray}
\mathcal{M}_0  &=&  6.76 \times 10^{10} \nonumber \\
f_0  &=&   0.336 \nonumber \\
\alpha  &=&   0.684 \\
\beta  &=&   0.205 \nonumber \\
\gamma  &=&   3.40\nonumber  
\end{eqnarray}
The green shaded region indicates the scatter about the best fit line,
which is 0.151 {\bf dex} for haloes with mass in the range of  
$3\times 10^{9}\Msun < M_{200} <2\times 10^{10} \Msun$, 0.236 for
halo mass in 
$2\times 10^{10}\Msun < M_{200} < 7\times 10^{10} \Msun$,
0.125 for halo mass in
$7\times 10^{10}\Msun < M_{200} < 3\times 10^{11} \Msun$
and 0.0518 for halo mass in
$3\times 10^{11}\Msun < M_{200} < 3.5\times 10^{12} \Msun$.

As might be expected, haloes with the highest masses we study have
high baryon fractions with relatively  small scatter. As halo mass
decreases, the baryon fraction decreases and the scatter increases,
consistent with expectations from outflows driven by stellar
feedback {\bf since the energy-driven wind model determined that
mass loading factors show a power-law dependency on circular velocity
, with an exponent of $\approx -2$, \eg \citet{Christensen16}, 
which means the greater efficiency of small galaxies at 
driving outflows is simply a result of their shallower potential 
well and the increasing scatters come from the increasing scatters of
stellar mass halo mass relation in lower mass range.}
At the lowest halo masses we study, below $10^{10}\Msun$,
the scatter starts to decrease. By contrast the scatter in the stellar
mass fraction increases below this scale. Thus it seems unlikely that
stellar feedback is primarily responsible for the low baryon
fractions.  Rather, we suggest an increased importance of the UV
background, which heats gas to above the virial temperature, thus
preventing it from collapsing into the low mass haloes. 
{\bf Although the three most massive galaixes in NIHAO suite have fairly 
high stellar masses (see blue filled circles in Fig.~\ref{fig:budget}),
they don't have large deviations apart from the fitting baryon fraction, 
and in the comparason with FIRE (see Fig.~\ref{fig:fire_com}), these
three objects seem to hint the trend on massive galaxy end associated
with the data from \citet{Voort16}. Moreover, the fitting relation 
about the baryon fraction is not sensitive to these three most massive 
galaxies as well since they have equal weight with other more than 
80 galaxies to shape the fitting curve.}

{\bf Since the resolution is one key parameter in simulation, and one
can imagine that with increased resolution, higher densities can be 
achieved in the CGM, potential leading to more cooling and larger
cold/cool gas fraction,
the convergence of the baryonic fraction predicted by NIHAO suite
is worth to estimate. Fig.~\ref{fig:budget_res} shows cold and cool
gas (HI + HeI) fraction color coded by number of dark matter particles 
per halo. In the plot, the number of dark matter partilce per halo spans 
2 magnitudes and has no correlation with cold and cool gas fraction. 
Therefore, we can conclude the resolution of NIHAO suite is 
proper to study the fractional baryon content.}

\subsection{Mass budget of the corona}
\label{sec:corona}

%% FIGURE 2
\begin{figure*}
\centerline{
  \psfig{figure=budgetpro.eps,width=1.0\textwidth}
}
\caption{Radial profile of the mass fraction of the gas in each phase
         to total baryonic mass in each radial bin 
         at $z=0$ for all galaxies in NIHAO sample.
         Each solid line is from one galaxy and colour coded with
         the halo mass.}
\label{fig:corona}
\end{figure*}

In Fig.~\ref{fig:corona}, we present the radial distribution of
gas in different phases at $z=0$, normalized to the total baryonic
mass profile, such that, in the region far away from central galaxy
where the stars are rare, for a given halo the four phases add up to
unity, {\bf each line is colour coded to the virial mass}.  
All simulations share a common attribute.  The cold gas (T$<
10^4$ K) is mostly located  near the center  ($R < 0.2 R_{200}$)
where most stars in galaxies form.  In contrast, the cool ($10^4$K
$<$ T $<$ $10^5$K) and  warm-hot ($10^5$K $<$ T $<$ $10^7$K) gas are
located at large distances with roughly constant fractions up to 2
times $R_{200}$.  The hot gas (T$>10^7$K) is a minority component
for all galaxies in the NIHAO sample, with the maximum hot gas
fraction at any radius being  less than 5\%.

Despite these similarities, we find a considerably higher proportion
of cool gas in lower mass galaxies (M$_{200} < 10^{11} \Msun$)
in the whole corona region.  For higher mass galaxies, warm-hot gas
dominates the corona  which signals {\bf stronger virial shock} and
higher efficiency  of feedback.
Even beyond the virial radius, the cool and warm-hot gas has similar
features as the gas within virial radius which reveals the gas
surrounding galaxies within large distance is the major reservoir of
baryons.

\subsection{Where are the missing baryons?}
\label{sec:where}

Fig.~\ref{fig:rps} shows the {\bf cumulative fraction} of total baryons  for
each simulation. Here the y-axis is the ratio between the baryonic to
dark matter mass, $M_{\rm b}(<R) / M_{\rm dm}(<R)$, enclosed
within a sphere of radius, $R$, normalized by the cosmic
baryon-to-dark matter ratio, $\Omega_{\rm b}/\Omega_{\rm dm}$.

Each solid curve represents a halo, and the curves are coloured by
their halo mass (red for high masses to blue for low masses).  Broadly
speaking, the curves have a similar shape, with a normalization that
depends on halo mass. They have a cusp in  the central region where the
stars and cold gas dominate, then become flat in the outer region.
More massive haloes have higher baryon fractions at all radii.  At
small radii, the baryon to dark matter ratio is higher than the cosmic
value due to gas dissipation.  Even beyond the virial radius, there is
little change in the baryon fraction up to 2 virial radii.  We thus
conclude that the missing baryons are well outside of the virial
radius.

{\bf Two cosmological hydrodynamic simulations in \citet{Ford15} 
which employed two prescriptions for galactic 
outflow models to compare with COS-Halos showed this cumulative
baryon fraction profile for COS-Halos matched sample of galaxies.
One hybrid energy/momentum driven winds model ("ezw" model) makes up to
65\% of all baryons inside the halo. This is broadly consistent with
estimates of baryonic mass derived from \citet{Werk14} and our 
finding in Fig.~\ref{fig:budget}. Comparing the cumulative baryon
fraction profile in detail, Fig.10 in \citet{Ford15} shows the baryon
fraction within 0.1$\Rvir$ is only $\sim 30\%$ and the fraction 
gradually increases to 65\% at $\Rvir$ while the profile of galaxies
with halo mass below $10^{11} \Msun$ in Fig.~\ref{fig:rps} have 
similar feature, the more massive galaxies all have roughly flat slopes.
The simplified outflow model (constant wind model or "cw" model) shows
lower fraction at all radii inside $\Rvir$ even though "ezw" model
and "cw" model generally gives similar observational absorption 
line properties. This suggests that the profile of cumulative baryon
fraction, as a complementary of total amount observation of CGM,
 is important to distinguish between competing outflow models
that often closely relate to the feedback models.
}
%% FIGURE 3
\begin{figure}
\centerline{
  \psfig{figure=ratioprofile.eps,width=0.5\textwidth}
}
\caption{Baryon distribution of each galaxy from NIHAO simulations. 
 The lines are colour coded by their halo mass, which shows a
  clear trend that the more massive haloes preserve more baryons at
  all radii.}
\label{fig:rps}
\end{figure}

To estimate how far the baryons escape, we measured the radius,
$\Rbar$, within which the total baryon mass equals to the fiducial
baryonic mass defined by Eq.~\ref{equ:mb}. This is a lower limit to
the true extent of the missing baryons since the baryon mass
includes gas and stars that belong to nearby lower mass haloes.
Fig.~\ref{fig:missvr} shows the baryon radius of each galaxy as
function of the virial mass.  We find the baryon radius generally
increases with virial mass. {\bf when normalized by the virial
radius, the distance baryons are ejected span a similar range of radii
$\Rbar/R_{200}\sim 2-6$ and gradually decrease as halo mass 
rise, even the majority of massive galaxies have $\Rbar \sim \Rvir$.}

%% FIGURE 4
\begin{figure}
\centerline{
  \psfig{figure=missing_vr.eps,width=0.5\textwidth}
}
\caption{Normalized baryon radius as function of total virial mass.
         The points are colour coded by  
         baryon radius of each galaxy.}
\label{fig:missvr}
\end{figure}

%% FIGURE 5
\begin{figure*}
\centerline{
  \psfig{figure=budget_inout.eps,width=1.0\textwidth}
}
\caption{Mass fraction of gas in four phases (relative to the fiducial
  baryonic mass within the virial radius)
  {\bf gas inside the virial radius (filled blue points) and between 
   virial and baryon radius (open red points)}, respectively. Cool gas is
  the dominant component of {\bf the fiducial baryonic mass within the 
  virial radius for most galaxies ($10^{10} < M/\Msun < 10^{12}$), cold 
   and warm-hot gas is majority only for the galaxies at lower and higher
   mass end.}}
\label{fig:inout}
\end{figure*}

In Fig.~\ref{fig:inout}, we show the mass fractions of gas (in four
phases) inside the virial radius (filled blue points) and between the
virial and baryon radius (open red points).
%
The fractions of cold gas,
$f_{\rm cold}$, are shown in the {\bf left} panel.  Inside the
virial radius $f_{\rm cold}$ increases gradually from zero at a halo
mass of $10^{10}\Msun$ to $\sim 20\%$ at a halo mass of
$10^{12}\Msun$.  The gas outside the virial radius has the opposite
and much stronger trend: $f_{\rm cold}\sim 50\%$ in haloes of mass
$10^{10}\Msun$ and decreases to zero by halo masses of
$10^{11}\Msun$.
%
The fractions of cool gas, $f_{\rm cool}$, are
shown in the {\bf middle} panel.  Inside the virial radius, $f_{\rm
  cool}$, has a maximum of 40\% at a halo mass of $10^{11} \Msun$,
and declines to less than 10\% below and above halo masses of
$10^{10}\Msun$ and $10^{12}\Msun$, respectively.  For most haloes there
is more cool gas outside than inside the virial radius.  In haloes
of mass $10^{10}\Msun$, $f_{\rm cool}\sim 80\%$, and decreases to
less than 10\% by a halo mass of $10^{12}\Msun$.
%
The fractions for warm-hot gas,$f_{\rm warm}$,  are shown in the
{\bf right} panel.  The trends of the gas inside and outside the virial
radius are quite similar, $f_{\rm warm}$ increases monotonically with
halo mass with maximum values of $\sim 30\%$.
%
The hot gas {\bf isn't shown} since it is negligible both
inside and outside the virial radius across the whole mass range we
study.
%
We thus conclude that, for galaxies with halo masses in the range
$10^{10} \lta M_{200}\lta 10^{11} \Msun$, the majority of baryons
associated with the dark matter halo are in the cold and cool phases,
and are located well outside of virial radius.  For haloes in the mass
range $10^{11} \lta M_{200}\lta 10^{12} \Msun$, the fractions of cold
gas, cool gas and  warm-hot gas are comparable.

\subsection{Comparison with observations
of Milky Way mass haloes}

{\bf Observations can gain information of CGM from absorption lines
and emission lines. Although emission lines allow us to direct obtain
a 3D picture of the distribution of gas in the CGM, emission line studies
preferentially probe the dense gas closer to galaxies since gas 
emissivity scales with the square of density. The situation is changing
by building current facilities, e.g.~\citet{Hayes16}, absorption line,
however, is the most common observational constraints on the physical
state of the CGM.}
The COS-HALOs survey is
filling in details about the $z \sim 0$ CGM \citep{Peeples14,
  Tumlinson11, Tumlinson13, Werk12, Werk13, Werk14}.  For the CGM of
low-redshift $\sim L^*$ galaxies ($\Mstar\sim 10^{10.5}\Msun$),
\citet{Tumlinson13} and \citet{Peeples14} constrain the mass of  the
warm-hot CGM ($T \sim 10^{5-7}$K), \citet{Werk14} provides a strict
lower limit to the mass of cool material ($T \sim 10^{4-5}$K) in the
CGM of these galaxies.  In a study using X-rays,  \citet{Anderson13}
place a constraints on the mass of  hot gas ($T > 10^7$K) residing in
the extended hot halos.

In Fig.~\ref{fig:comparison}, we show the mean values and standard deviation
of the mass fraction of stars and different components of gas in our
most massive galaxies ($3.49 \times 10^{11} \Msun$ $<$ M$_{200}$ $<$ 
$3.53 \times 10^{12} \Msun$) with blue points and error bars.
The gas is assigned to a range of temperature bins: 
cold gas (T $< 10^4$ K), cool gas ($10^4$ K $<$ T $< 10^5$ K), 
warm gas ($10^5$ K $<$ T $< 10^7$ K) and hot gas (T $> 10^7$ K).
The observational constraints are shown with the same colour scheme
in Fig. 11 in \citet{Werk14}.

%% FIGURE 6
\begin{figure}
\centerline{
  \psfig{figure=budget.eps,width=0.5\textwidth}
}
\caption{Baryonic budget 
  of NIHAO haloes of mass $3.5\times 10^{11} < M_{200}/\Msun <
  3.5 \times 10^{12}$ (blue points 
  with 1$\sigma$ error bars) compared with observations of 
  $M_{200} \sim 10^{12}\Msun$ haloes (shaded regions). There is good agreement, except for the cool gas which has two conflicting measurements.
  }
\label{fig:comparison}
\end{figure}

%% TABLE 1
\begin{table*}
  \caption{The baryonic budget parameters for NIHAO galaxies
    in different halo  mass bins. We refer to gas in the temperature range  T
  $<$ $10^4$ K as cold; $10^4$ K $\leqslant$ T $<$ $10^5$ K as cool;
  $10^5$ K $\leqslant$ T $<$ $10^7$ K as warm;  and T $\geqslant$
  $10^7$ K as hot.}
\begin{center}
\begin{tabular}{ccccc}
\hline
\input{budgettable4.txt}
\hline
\end{tabular}
\label{tab:comparison}
\end{center}
\end{table*}

In this plot \citet{Werk14} provides observational constraints for
CGM gas mass that are shown as the shaded bars.  The stellar  mass
range comes from halo abundance matching as described in
\citet{Kravtsov14}.   The cold disk gas mass comes from from
\citet{Dutton11}.

The observations and the simulations match well in every phase except
 the cool CGM gas, where the observations find $3\times$ the
mass that simulations predict.  If the observations are correct, the
simulations have either ejected cool gas too far, or  they have
created a CGM with the wrong mix of gas temperatures. The total gas
fractions (0.39 in COS-HALOs, 0.41 in NIHAO) suggest the latter
option.   However, \citet{Stern16} developed a new method to
constrain the physical conditions in the cool CGM from measurements
of ionic  columns densities. This new method combines the
information available from different sightlines during the photoionization
modeling, and was applied to the COS-HALOs data,  yielding a total
cool CGM mass within the virial radius of $1.3\times10^{10}\Msun$
which is shown by the green hashed bar in Fig~\ref{fig:budget} and
is in good agreement with our prediction.  As the
Fig.~\ref{fig:corona} and Fig.~\ref{fig:inout} show, the cool gas is
the most important component {\bf for most galaxies in NIHAO simulation
($10^{10} < M/\Msun < 10^{12}$), the cold and hot gas only dominate for
galaxies at lower and higher mass end in NIHAO suite.} 
{\bf A new estimation for the baryonic fractions from a set of
eight Milky Way-sized zoom-in cosmological simulations 
\citep{Colin16} is genenrally consistence with our results. The 
fraction of cool gas from their simulation is even lower 
($f_{\rm cool}$ = 0.034) than ours and the result from 
\citet{Stern16}. On the other hand, \citet{Peeples14} gave 
a conservative observational estimate about warm-hot gas and
showed that the fraction of gas in this phase is only 5\%, 
which is the lower limit of the estimates from \citet{Werk14} 
and less than the average values from \citet{Colin16} 
($f_{\rm warm-hot} \approx 24\%$) and our finding 
($f_{\rm warm-hot} \approx 17\%$). Therefore, the mix temperature
of CGM have large uncertainty and}
the more accurate
knowledge of the physical properties of CGM are necessary to better
understand the role of the CGM in galaxy formation.
%
As the CGM of lower mass galaxies will soon be observed,
Table~\ref{tab:comparison} lists information about CGM
mass fractions of the different components of gas in haloes down to a
halo mass of $\sim 10^{10}\Msun$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% SECTION 4: SUMMARY
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Summary}
\label{sec:sum}

We have used the NIHAO galaxy simulation suite \citep{Wang15} to study
the statistical features of the baryonic budget and distribution
spanning halo masses of $\sim 10^{10}$ to $\sim 10^{12}\Msun$. NIHAO
is a large (currently 88) set of high resolution cosmological
hydrodynamical galaxy formation simulations. As shown in previous papers
the NIHAO galaxies reproduce several key observed scaling relations.
We summarize our results as follows:

\begin{itemize}
\item All of the NIHAO haloes have a lower baryon to dark matter ratio, 
      inside the virial radius, than the cosmic baryon fraction. 
      {\bf Higher fractions of baryons are lost as the virial mass 
      decreases in the mass range of $\lesssim 10^{12} \Msun$.
      The trend confirmed with \citet{Christensen16} and \citet{Voort16},
      and associated by \citet{Voort16}, the trend at higher mass range
      may suggest the baryon fraction is reduced larger as virial mass
      increase. }

\item Lower mass haloes have lost a larger fraction of their baryons, 
      even though they convert a lower fraction of the baryons into stars.

\item Cold gas ($T<10^4$K) is mostly restricted to be within
      0.2 virial radii.  The cool gas ($10^4 < T < 10^5$K) dominates
      the corona at low masses ($M_{200}\lta 3\times 10^{11} \Msun$)
      while the warm-hot gas ($10^5 < T < 10^7$K) dominates at high
      masses ($M_{200}\gta 3\times 10^{11} \Msun$).

\item {\bf All haloes have less baryons than baryon mass derived by cosmic
       baryonic fraction even beyond 2 virial radii. The total 
       baryons derived by cosmic baryonic fraction within virial radius 
       can expand more than $8\Rvir$ for low mass haloes but less than
       $2\Rvir$ for high mass haloes.}

\item Cool gas is a major component of the total baryons within
      R$_{\rm bar}$ {\bf for galaxies at mass range between $10^{10} \Msun$ 
      to $10^{12} \Msun$ cold and warm-hot gas dominates only for galaxies 
      at lower and higher mass end. Most of baryons are} outside of the 
      virial radius.

\item For the highest mass haloes in our study $\sim 10^{12}\Msun$ our
      simulations are consistent with the observed  fractions of stars,
      cold gas, warm and hot gas \citep{Werk14}. {\bf \citet{Peeples14} 
      showed a conservative observational estimate for warm-hot gas 
      which is lower limit of the result from \citet{Werk14} and the 
      warm-hot fraction from our study is well within the range of less
      conservative estimates.}

\item For the cool gas we predict $f_{\rm cool}=0.11\pm0.06$
      which is significantly lower than the observations from COS-HALOs
      ($f_{\rm cool}=0.28-0.48$), but is in excellent agreement with the
      analysis of \citet{Stern16}.
      {\bf As in \citet{Colin16}, the simulations suggest that the
      fraction of cool gas is approximately 10\%.}

\end{itemize}

\section*{Acknowledgments}

We thank the two anonymous referees whose suggestions greatly 
improve the paper.
{\bf We thank Freeke van de Voort for kindly sharing data of the FIRE
simulation.}
{\sc Gasoline} was written by Tom Quinn and James Wadsley. Without
their contribution, this paper would have been impossible.
%
The simulations were performed on the {\sc theo} cluster of the
Max-Planck-Institut f\"ur Astronomie and the {\sc hydra} cluster at
the Rechenzentrum in Garching; and the Milky Way supercomputer, funded
by the Deutsche Forschungsgemeinschaft (DFG) through Collaborative
Research Center (SFB 881) "The Milky Way System" (subproject Z2),
hosted and co-funded by the J\"ulich Supercomputing Center (JSC). We
greatly appreciate the contributions of all these computing
allocations.
%
AAD, GSS and AVM acknowledge support through the
Sonderforschungsbereich SFB 881 “The Milky Way System” (subproject A1)
of the German Research Foundation (DFG).  The analysis made use of the
pynbody package \citep{Pontzen13}.
%
The authors acknowledge support from the MPG-CAS through the
partnership programme between the MPIA group lead by AVM and the PMO
group lead by XK.
%
LW acknowledges support of the MPG-CAS student programme.
%
XK acknowledge the support from 973 program (No. 2015CB857003,
2013CB834900), NSFC project No.11333008 and the ``Strategic Priority
Research Program the Emergence of Cosmological Structures'' of the
CAS(No.XD09010000).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  REFERENCES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{thebibliography}{}

%%AAAAAAAA

% Extended Hot Halos around Isolated Galaxies Observed in the ROSAT All-Sky Survey
\bibitem[Anderson et al.(2013)]{Anderson13} Anderson, M.~E., Bregman, J.~N., Dai, X.\ 2013, \apj, 762, 106

% Fundamental differences between SPH and grid methods
\bibitem[Agertz et al.(2007)]{Agertz07} Agertz, O., Moore, B., Stadel, J.\ 2007, \mnras, 380, 963

%%BBBBBBBB

%The Average Star Formation Histories of Galaxies in Dark Matter Halos
% from z = 0-8
\bibitem[Behroozi et al.(2013)]{Behroozi13} Behroozi, P.~S.,
  Wechsler, R.~H., \& Conroy, C.\ 2013, \apj, 770, 57

%A First Estimate of the Baryonic Mass Function of Galaxies
\bibitem[Bell et al.(2003)]{Bell03} Bell, E.~F., McIntosh, D.~H., Katz, N., Weinberg, M.~D.,\ 2003, \apj, 585, 117

% The Search for the Missing Baryons at Low Redshift
\bibitem[Bregman (2007)]{Bregman07}
Bregman, J.~N.\ 2007, ARAA, 45, 221

%% CCCCCCCC

%Where are the baryons
\bibitem[Cen \& Ostriker (1999)]{Cen09} 
Cen, R.~Y., Ostriker, J.~P.\ 1999, \apj, 514, 1

%In-N-Out: The Gas Cycle from Dwarfs to Spiral Galaxies
\bibitem[Christensen et al.(2016)]{Christensen16}
Christensen, C.~R., Dav{\'e}, R., Governato, F., et al.\ 2016, \apj, 824, 57

%Cosmological simulations of Milky Way-sized galaxies
\bibitem[Colin et al.(2016)]{Colin16}
Colin, P., Avila-Reese, V., Roca-Fabrega, S., et al.\ 2016, \apj, in press

%% DDDDDDDD

% Missing Halo Baryons and Galactic Outflows
\bibitem[Dav{\'e} (2009)]{Dave09} Dav{\'e}, R.\ 2009, ASPC, 419, 347D

% The intergalactic medium over the last 10 billion years - I. Lya absorption and physical conditions
\bibitem[Dav{\'e} et al.(2010)]{Dave10} Dav{\'e}, R., Oppenheimer, B.~D., Katz, N., et al.\ 2010, \mnras, 408, 2051

%Dark halo response and the stellar initial mass function in early-type and late-type galaxies
\bibitem[Dutton et al.(2011)]{Dutton11} Dutton, A.~A., Conroy, 
  C., van den Bosch, F.~C., et al.\ 2011, \mnras, 416, 322

%The baryonic Tully-Fisher relation and galactic outflows
\bibitem[Dutton(2012)]{Dutton12} Dutton, A.~A.\ 2012,
  \mnras, 424, 3123

%Cold dark matter haloes in the Planck era: evolution of structural parameters for Einasto and NFW profiles
\bibitem[Dutton \& Macci{\`o}(2014)]{Dutton14} Dutton,
  A.~A., \& Macci{\`o}, A.~V.\ 2014, \mnras, 441, 3359

%% EEEEEEEEE

%% FFFFFFFFF

%
\bibitem[Ford et al.(2013)]{Ford13} Ford, A.~B., Oppenheimer, B.~D., Dav{\`e}, R., et al.\ 2013, \mnras, 432, 89

%
\bibitem[Ford et al.(2015)]{Ford15} Ford, A.~B., Werk, J.~W., Dav{\`e}, R., et al.\ 2016, \mnras, 459, 1745

% The Cosmic Baryon Budget
\bibitem[Fukugita et al.(1998)]{Fukugita98} Fukugita, M., Hogan, C.~J., Peebles, P.~J.~F.\ 1998, \apj, 503, 518

%% GGGGGGGGG

% NIHAO VIII:
\bibitem[Gutcke et al.(2016)]{Gutcke16} Gutcke, T.~A., Stinson, G.~S., Macci{\`o}, A.~V., et al.\ 2016, arXiv:1602.06956

%% HHHHHHHHH

% Large-scale mass distribution in the Illustris simulation
\bibitem[Haider et al.(2016)]{Haider16}
Haider, M., Steinhauser, D., Vogelsberger, M., et al.\ 2016, \mnras, 457, 3024

% O VI Emission Imaging of a Galaxy with the Hubble Space Telescope: a Warm Gas Halo Surrounding the Intense Starburst SDSS J115630.63+500822.1O VI Emission Imaging of a Galaxy with the Hubble Space Telescope: a Warm Gas Halo Surrounding the Intense Starburst SDSS J115630.63+500822.1
\bibitem[Hayes et al.(2016)]{Hayes16}
Hayes, M., Melinder, J., {\"O}stlin, G., et al.\ 2016, arXiv:1606.04536

% Distributions of the Baryon Fraction on Large Scales in the Universe
\bibitem[He et al.(2005)]{He05}
He, P., Feng, L.~L., Fang, L.~Z.\ 2005, \apj, 623, 601

% Galaxies on FIRE (Feedback In Realistic Environments): stellar
% feedback explains cosmologically inefficient star formation
\bibitem[Hopkins et al.(2014)]{Hopkins14} Hopkins, P.~F., Kere{\v s}, D., O{\~n}orbe, J., et al.\ 2014, \mnras, 445, 581

%
\bibitem[Hummels et al.(2013)]{Hummels13} Hummels, C.~B., Bryan, G.~L., Smith, B.~D., et al.\ 2013, \mnras, 430, 1548

%% JJJJJJJJJJJ

%% KKKKKKKKKKK

%A superbubble feedback model for galaxy simulations
\bibitem[Keller et al.(2014)]{Keller14} Keller, B.~W., Wadsley, 
  J., Benincasa, S.~M., \& Couchman, H.~M.~P.\ 2014, \mnras, 442, 3013

% Stellar mass -- halo mass relation and star formation efficiency in
% high-mass halos
\bibitem[Kravtsov et al.(2014)]{Kravtsov14} Kravtsov, A., 
Vikhlinin, A., \& Meshscheryakov, A.\ 2014, arXiv:1401.7329

%% LLLLLLLLL

%% MMMMMMMMM

% On the origin of the warm-hot absorbers in the Milky Way's halo
\bibitem[Marasco et al.(2013)]{Marasco13} 
Marasco, A., Marinacci, F., Fraternali, F.\ 2013, \mnras, 433, 1634

% The Baryon Content of Cosmic Structures
\bibitem[McGaugh et al.(2010)]{McGaugh10} McGaugh, S.~S., 
Schombert, J.~M., de Blok, W.~J.~G., Zagursky, M.~J.\ 2010, \mnras,
708, 14

% Galactic star formation and accretion histories from matching galaxies to dark matter haloes
\bibitem[Moster et al.(2013)]{Moster13} Moster, B.~P., Naab, T., 
\& White, S.~D.~M.\ 2013, \mnras, 428, 3121

% Gusty, gaseous flows of FIRE: galactic winds in cosmological simulations with explicit stellar feedback
\bibitem[Muratov et al.(2015)]{Muratov15} Muratov, A.~L., Ker{\v e}s, D., Faucher-Gigu{\`e}re, C., et al.\ 2015, \mnras, 454, 2691

%% NNNNNNNNN

%% OOOOOOOO

%
\bibitem[Oppenheimer et al.(2016)]{Oppenheimer16} Oppenheimer, B.~D., Crain, R.~A., Schaye, J., et al.\ 2016, arXiv:1603.05984

%% PPPPPPPPP

%Planck 2013 results. XVI. Cosmological parameters
\bibitem[the Planck Collaboration et  al.(2014)]{Planck14}
  Planck Collaboration, Ade, P.~A.~R., Aghanim, N., et al.\ 2014,
  \aap, 571, AA16

% A Budget and Accounting of Metals at z ~ 0: Results from the COS-Halos Survey
\bibitem[Peeples et al.(2014)]{Peeples14} Peeples, M.~S., Werk, J.~K., Tumlinson, J., et al.\ 2014, \apj, 786, 54

% The baryon content of the universe
\bibitem[Persic \& Salucci(1992)]{Persic92} Persic, M., Salucci, P.\ 1992, \mnras, 258, 14

%pynbody: N-Body/SPH analysis for python
\bibitem[Pontzen et al.(2013)]{Pontzen13} Pontzen, A., Ro{\v s}kar, R., Stinson, G., \& Woods, R.\ 2013, Astrophysics Source Code Library, 1305.002

%% RRRRRRRRRRR

%% SSSSSSSSSSSS

% The EAGLE project: simulating the evolution and assembly of
% galaxies and their environments
\bibitem[Schaye et al.(2015)]{Schaye15} Schaye, J., Crain,
  R.~A., Bower, R.~G., et al.\ 2015, \mnras, 446, 521

% The enrichment of the intergalactic medium with adiabatic feedback - I. Metal cooling and metal diffusion
\bibitem[Shen et al.(2010)]{Shen10} Shen, S., Wadsley, J., 
\& Stinson, G.\ 2010, \mnras, 407, 1581

% On the structure of hot gas in haloes: implications for the LX-TX relation and missing baryons
\bibitem[Sharma et al.(2012)]{Sharma12}
Sharma, P., McCourt, M., Parrish, I.~J., Quataert, E.\ 2012, \mnras, 427, 1219

% The Baryon Census in a Multiphase Intergalactic Medium: 30% of the Baryons May Still be Missing
\bibitem[Shull et al.(2012)]{Shull12}
Shull, J.~M., Smith, B.~D., Danforth, C.~W.\ 2012, \apj, 759, 23

%
\bibitem[Shull (2014)]{Shull14} Shull, J.~M.\ 2014, \apj, 784, 142

% Diffuse Coronae in Cosmological Simulations of Milky Way-sized Galaxies
\bibitem[Sokolowska et al.(2016)]{Sokolowska16}
Sokolowska, A., Mayer, L., Babul, A., Madau, P., Shen, S.\ 2016, \apj, 819, 21

% A Universal Density Structure for Circum-Galactic Gas
\bibitem[Stern et al.(2016)]{Stern16} 
Stern, J., Hennawi, J.~F., Prochaska, J.~X., Werk, J.~K.\ 2016, arXiv:1604.02168

% Star formation and feedback in smoothed particle hydrodynamic simulations - I. Isolated galaxies
\bibitem[Stinson et al.(2006)]{Stinson06} Stinson, G.~S., Seth, A., Katz, N., et al.\ 2006, \mnras, 373, 1074

%
\bibitem[Stinson et al.(2012)]{Stinson12} Stinson, G.~S., Brook, C., Prochaska, J.~X., et al.\ 2012, \mnras, 425, 129

%Making Galaxies In a Cosmological Context: the need for early stellar feedback
\bibitem[Stinson et al.(2013)]{Stinson13} Stinson, G.~S., Brook, 
C., Macci{\`o}, A.~V., et al.\ 2013, \mnras, 428, 129

% NIHAO III: the constant disc gas mass conspiracy
\bibitem[Stinson et al.(2015)]{Stinson15} Stinson, G.~S., Dutton, A.~A., Wang, L., et al.\ 2015, \mnras, 454, 1105

%
\bibitem[Suresh et al.(2015)]{Suresh15b} Suresh, J., Rubin, K.~H.~R., Kannan, R., et al.\ 2015, arXiv:1511.00687

%% TTTTTTTTTTT

% Not Dead Yet: Cool Circumgalactic Gas in the Halos of Early-type Galaxies
\bibitem[Thom et al.(2012)]{Thom12}
Thom, C., Tumlinson, J., Werk, J.~K.\ 2012, \apjl, 758, L41

% The Large, Oxygen-Rich Halos of Star-Forming Galaxies Are a Major Reservoir
% of Galactic Metals
\bibitem[Tumlinson et al.(2011)]{Tumlinson11} Tumlinson, J., Thom, C., Werk, J., et al.\ 2011, Science, 334, 948

% The COS-Halos Survey: Rationale, Design and a Census of Circumgalactic Neutral Hydrogen
\bibitem[Tumlinson et al.(2013)]{Tumlinson13} Tumlinson, J., Thom, C., Werk, J., et al.\ 2013, \apj, 777, 59

%% VVVVVVVVVVVV

% Properties of gas in and around galaxy haloes
\bibitem[van de Voort \& Schaye (2012)]{Voort12} van de Voort, F., Schaye, J.\ 2012, \mnras, 423, 2991

%The impact of stellar feedback on hot gas in galaxy haloes:
%the Sunyaev-Zeldovich effect and soft X-ray emission 
\bibitem[van de Voort et al.(2016)]{Voort16} van de Voort, F., Quataert, E., Hopkins, P.~F., et al.\ 2016, arXiv:1604.01397

%	Introducing the Illustris Project: simulating the coevolution
% of dark and visible matter in the Universe
\bibitem[Vogelsberger et al.(2014)]{Vogelsberger14} Vogelsberger, M., 
Genel, S., Springel, V., et al.\ 2014, \mnras, 444, 1518

%% WWWWWWWWWWWWWW

%Gasoline: a flexible, parallel implementation of TreeSPH
\bibitem[Wadsley et al.(2004)]{Wadsley04} Wadsley, J.~W., Stadel, 
J., \& Quinn, T.\ 2004, \na, 9, 137

%On the treatment of entropy mixing in numerical cosmology
\bibitem[Wadsley et al.(2008)]{Wadsley08} Wadsley, J.~W., 
Veeravalli, G., \& Couchman, H.~M.~P.\ 2008, \mnras, 387, 427

  %NIHAO project I: Reproducing the inefficiency of galaxy formation across cosmic time with a large sample of cosmological hydrodynamical simulations
\bibitem[Wang et al.(2015)]{Wang15} Wang, L., Dutton, A.~A.,  Stinson, G.~S., et al.\ 2015, \mnras, 454, 83

%  A fundamental problem in our understanding of low-mass galaxy evolution
  \bibitem[Weinmann et al.(2012)]{Weinmann12} Weinmann, S.~M., Pasquali, A., Oppenheimer, B.~D., et al.\ 2012, \mnras, 426, 2797

% The COS-Halos Survey: Keck LRIS and Magellan MagE Optical Spectroscopy
\bibitem[Werk et al.(2012)]{Werk12} Werk, J.~k., Prochaska, J.~X., Thom, C., et al.\ 2012, \apjs, 198, 3

% The COS-Halos Survey: An Empirical Description of Metal-line Absorption in the Low-redshift Circumgalactic Medium
\bibitem[Werk et al.(2013)]{Werk13} Werk, J.~k., Prochaska, J.~X., Thom, C., et al.\ 2013, \apjs, 204, 17

%The COS-Halos Survey: Physical Conditions and Baryonic Mass in the Low-redshift Circumgalactic Medium
\bibitem[Werk et al.(2014)]{Werk14} Werk, J.~k., Prochaska, J.~X., Thom, C., et al.\ 2014, \apj, 792, 8

%% YYYYYYYYYYYY

% The Temperature Structure of the Warm-Hot Intergalactic Medium
\bibitem[Yoshida et al.(2005)]{Yoshida05} 
Yoshida, N., Furlanetto, S.~R., Hernquist, L.\ 2005, \apj, 618L, 91

%% ZZZZZZZZZZZZZ

% Dynamical effect of the turbulence of the intergalactic medium on the baryon fraction distribution
\bibitem[Zhu et al.(2011)]{Zhu11} Zhu, W., Feng, L.~L., Fang, L.~Z.\ 2011, \mnras, 415, 1093

\end{thebibliography}

\label{lastpage}
\end{document}